---
title: "Olympic Games"
author: Sayra Moore, Andrea Piolini, Sabina Shahbazzade, Caroline Sklaver, Spencer
  Staub
date: "10/13/2019"
output:
  html_document:
    number_sectiond: yes
    toc: yes
    toc_float:
      collapsed: no
  word_document:
    toc: yes
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
```
 Olympic Games

# Chapter 1: Data File

Olympic Games dataset is a historical dataset that contains data from Athens 1886 to Rio 2016. The dataset was found on Kaggle ( https://www.kaggle.com/heesoo37/120-years-of-olympic-history-athletes-and-results) and originally scrapped from sports-reference.com in May 2018. Our dataset consists of 27116 observations and 15 variables/columns that are ID (unique number of each athlete), Name (athlete’s name), Sex (M or F), Age (Integer), Weight (in kilograms), Height (in centimeters), Team (Team name), NOC (National Olympic Committee 3-letter code), Games (year and season), Year (integer), Season (Summer or Winter), City (Host city), Sport, Event,  Medal (gold, silver, bronze, or NA).

Several Kaggle users have conducted various analyses on this dataset using Python and R. Some of these analyses specifically focused on single events, other focused on the analysis of specific sports and some others focused on the variation of age, gender, height and weight over time. 

We chose this dataset because of our passion for sports and our shared excitement every time the Olympic Games take place. Therefore, we thought we could use our statistical and R expertise to learn more about these events.

When we looked at the dataset, we thought it would be very interesting to test whether certain characteristics such as height and age have an effect on whether an athlete earns a medal. We also decided to focus on this question because the majority of the analyses on the dataset focus more on descriptive statistic and data visualization, and they do not perform statistical tests to see whether the variables are related and correlated.

```{r read, echo=FALSE, include = TRUE}
#setwd('C:/Users/andre/Desktop/GW Classes/Introduction to Data Science')
athlete_events = read.csv("athlete_events.csv")
str(athlete_events)
```

# Chapter 2: Limitations of the Dataset

There are four limitations in our dataset: 

1.	The dataset is not complete. There are many NAs for Age, Height, Weight.

2.	It does not contain the country hosting the event, it has only the city of event.

3.	More information on the country hosting the event and on the ethnicity of the athletes would be beneficial 

4.	Women did not begin participating until the 1900 Paris Olympics


```{r NAs, echo=FALSE}
# omitting NAs leaves only athletes that won medals
summer_events = na.omit(athlete_events[athlete_events$Season == 'Summer',])

# keeping NAs for analysis of winners and losers
summer_w_NAs = athlete_events[athlete_events$Season=='Summer',]
```

# Chapter 3: Descriptive Statistics


For descriptive statistics we have the following questions for our dataset:

How many athletes participated over the years? How many males? How many females? How many medals were won by country? How many medals were won by males and females? Is there correlation between age, weight, and height? Do these have an effect on winning or not? Is there difference in average age between males and females? Does their average vary over years? 

## Chapter 3.1: Medals

The total number of Summer Olympic Games is 29 over 120 years (every 4 years except for 1916 when it was canceled due to WWI). The average total number of medals earned by each country is 36.72609.

```{r Descriptive Stats Medals, echo=FALSE, include=FALSE}
#Total Number of Summer Games
unique_games = unique(summer_w_NAs$Games)
length(unique_games)

# Medals per country
medals_table = table(summer_events$Medal,summer_events$NOC)
mean(medals_table)

```

```{r Descriptive Stats Medals Graphs, echo=FALSE, include=TRUE}
# stacked barplot of medals won by top countries
subset_medals = medals_table[,c('USA', 'URS', 'GBR', 'GER', 'ITA', 'FRA', 'AUS',
                                'CHN','HUN','JPN','NED','RUS','SWE')]
barplot(subset_medals, main = "Total medals won by country", 
        ylab = 'Country Code', xlab = "Number of medals won", col = c('gold', 'light gray','gold4'),
        legend.text = c("Gold",'Silver','Bronze'), horiz = TRUE, las=2)



# Graph of total numebr of medals won by men and women
athlete_events_MPYG = summer_events %>% select(Team, Medal, Sex, Year) %>% group_by(Team, Year, Sex)
medal_distribution_peryeargender = count(athlete_events_MPYG, Medal)
medal_distribution_peryeargender = subset(medal_distribution_peryeargender, medal_distribution_peryeargender$Medal != "NA")
ggplot(medal_distribution_peryeargender, aes(x = Sex, y = n, fill = Sex)) + geom_bar(stat = 'identity', width = 0.5) + scale_fill_manual(values=c('F' = 'deeppink1', 'M' = 'dodgerblue1')) + ggtitle('Comparison of Medals Won by Men and Women') + labs(x = 'Sex', y = 'Number of Medals')

```

Graphical representation of top countries total medals earned by medal type. This clearly shows teams that have dominated the 29 Olympic games. According the graph, we can see that the highest number of golds, silvers and bronze medals was won by USA (United States of America). The second highest number of golds, silvers and bronze medals was URS (Soviet Union). The lowest number of medals was won by SWE (Sweden). GER (Germany) and AUS (Australia) have almost the same numbers of all medals as well as GBR (Great Britain), CHN (China), and RUS (Russia). ITA (Italia) and FRA (France) have also the same number of medals. JPN (Japan) and HUN (Hungary) have the same number of golds, silvers, and bronze medals. 

It is also clear that men won more medals than women, where number of medals won by men more than 15,000 and women less than 10,000. One reason of less medals won by women was that women did not participated Olympic games before 1900. 

## Chapter 3.2: Athletes Participation Over 120 Years of Summer Olympic Games

Graphical representation of number of athletes over the 120 years of Olympics. Additionally, graph of male and female athlete participation over time to visualize the difference in participation by gender.


### Chapter 3.2.1: Participation by Year

```{r Descriptive Statistics Athletes, echo=FALSE, include=TRUE}
# participants per Year
library(dplyr)
athlete_events_PPYnn = summer_w_NAs %>% select(Name, Year) %>% group_by(Year)
participants_distribution = count(athlete_events_PPYnn, Year)
participants_distribution = subset(participants_distribution, participants_distribution$Year != "NA")

# participants over time
athlete_events_MPC = summer_w_NAs %>% group_by(Year) %>% summarise(count = n_distinct(Team))
ggplot(data=athlete_events_MPC, aes(x=Year, y=count)) +
  geom_line()+
  geom_point() + ggtitle('Number of Olympic athletes over time') + ylab('Number of participants')


# plot male female participants over time
group_count_mf <- summer_w_NAs %>% group_by(Year, Sex) %>%
  summarize(Athletes = length(unique(ID)))
group_count_mf$Year <- as.integer(group_count_mf$Year)

ggplot(group_count_mf, aes(x=Year, y=Athletes, group=Sex, color=Sex)) +
  geom_point(size=2) +
  geom_line()  +
  scale_color_manual(values=c("deeppink1","dodgerblue1")) +
  labs(title = "Number of male and female Olympic Athletes over time") +
  theme(plot.title = element_text(hjust = 0.5))

```

In our participation by year graphs we can see a general increase in participation from the inception of the Olympic games, however, there are a few points where participation decreased. We have hypothesized that these decreases in participation are largely due to wars that encompass a significant amount of the world (ie. World War 1, World War 2, Korean War). We can see that drops in population have affected men a significant amount more than females, which backs up our hypothesis.

### Chapter 3.2.2: Participation by Sport

```{r Descriptive Statistics Athletes 2, echo=FALSE, include=TRUE}
# participation by sport
sport_participation = sort(table(summer_w_NAs$Sport), decreasing = TRUE)
barplot(sport_participation[2:20], las=2, main="Total Atheletes by Sport", col = 'gray78')

# male and female by sport
female_w_NAs = summer_w_NAs[summer_w_NAs$Sex == 'F',]
male_w_NAs = summer_w_NAs[summer_w_NAs$Sex == 'M',]

female_sport_participation = sort(table(female_w_NAs$Sport), decreasing = TRUE)
male_sport_participation = sort(table(male_w_NAs$Sport), decreasing = TRUE)


barplot(female_sport_participation[2:10], las=2, main="Number of Female Athletes by Sport", col = 'deeppink1',
        ylim = c(0,18000))
barplot(male_sport_participation[2:10], las=2, main="Number of Male Athletes by Sport", col='dodgerblue1',
        ylim = c(0,18000))
```

We can also conclude from our participation distribution by sports graphs that gymnastics and swimming are by far the most popular sports of the Olympics, however, for females the two aforementioned sports categories make up a predominant amount of participation. For Males gymnastics and swimming are still quite popular, but the distribution of participation is significantly more even than their female counterparts.

## Chapter 3.3: Age, Height, Weight

```{r , echo=FALSE, include=TRUE}
# Descriptive stats of Age, Height, Weight

par(mfrow=c(1,3))

hist(summer_w_NAs$Age, breaks = 30, main = 'Histogram of Age', xlab = 'Age', ylab = 'Frequency', col = 'rosybrown1')
#hist(f$Age, breaks = 30)
#hist(m$Age, breaks = 30)

hist(summer_w_NAs$Height, breaks = 30, main = 'Histogram of Height',xlab = 'Height(cm)', ylab = 'Frequency', col = 'darkseagreen3')
#hist(f$Height, breaks = 30)
#hist(m$Height, breaks = 30)

hist(summer_w_NAs$Weight, breaks = 30, main = 'Histogram of Weight',xlab = 'Weight(kg)', ylab = 'Frequency', col = 'lightblue3')
#hist(f$Weight, breaks = 30)
#hist(m$Weight, breaks = 30)

```

Histograms to view the distribution of Age, Height, and Weight for all Olympic Athletes. This is useful to determine the distribution. Age and weight are a bit right-skewed, height is very normal. These distributions do not vary greatly when looking at males and females separately. 

```{r AHW Correlations, echo=FALSE, include=TRUE}
# Corrrelations between age, height, weight
corr_col = cor(na.omit(summer_w_NAs[c(4,5,6)]))
corr_col
library(corrplot)
corrplot(corr_col, method = "circle")

```

Correlations between age, height, and weight, also represented in graphics. Height and weight are highly correlated (0.7951830). 

# Chapter 4: Statistical Tests

Our inspiration for the following tests revolved around questions regarding medals won and participant statistics (age, height, weight), as well as participant statistics over time.

## Chapter 4.1: Does age, height, or weight determine outcome (winning of medals)?

Chi-squared tests in which the null hypothesis is that age, height, and weight do not have a significant effect on whether athletes won a medal or not.These tests resulted in p-values <0.05 meaning we can reject the null hypotheses and conclude that age, height, and weight all determine whether athletes win medals. The age, height, and weight data was made categorical into intervals based on quartiles and mean of each variable. 

```{r Chi-Squraed, echo=FALSE, include=TRUE}
# Chi Squared: Does age, height, or weight determine if you won a medal or not

# differenciate between winners and losers
summer_w_NAs <- subset(athlete_events, Season == 'Summer')
summer_w_NAs$Medal <- as.character(summer_w_NAs$Medal)
summer_w_NAs$Medal[is.na(summer_w_NAs$Medal)] <- "Loser"
summer_w_NAs$Medal[summer_w_NAs$Medal == 'Bronze'] <- 'Winner'
summer_w_NAs$Medal[summer_w_NAs$Medal == 'Silver'] <- 'Winner'
summer_w_NAs$Medal[summer_w_NAs$Medal == 'Gold'] <- 'Winner'

# make age/height/weight categorical
summary(summer_w_NAs$Age)
summer_w_NAs$Cat_Age = summer_w_NAs$Age
summer_w_NAs$Cat_Age = replace(summer_w_NAs$Cat_Age, summer_w_NAs$Cat_Age >= 28, 'old')
summer_w_NAs$Cat_Age = replace(summer_w_NAs$Cat_Age, summer_w_NAs$Cat_Age<=21, 'young')
summer_w_NAs$Cat_Age = replace(summer_w_NAs$Cat_Age, 
                               summer_w_NAs$Cat_Age > 21 & summer_w_NAs$Cat_Age < 28, 'avg')


summary(summer_w_NAs$Height)
summer_w_NAs$Cat_Height = summer_w_NAs$Height
summer_w_NAs$Cat_Height = replace(summer_w_NAs$Cat_Height, summer_w_NAs$Cat_Height >= 183, 'tall')
summer_w_NAs$Cat_Height = replace(summer_w_NAs$Cat_Height, summer_w_NAs$Cat_Height<=168, 'short')
summer_w_NAs$Cat_Height = replace(summer_w_NAs$Cat_Height, 
                               summer_w_NAs$Cat_Height > 168 & summer_w_NAs$Cat_Height < 183, 'avg')


summary(summer_w_NAs$Weight)
summer_w_NAs$Cat_Weight = summer_w_NAs$Weight
summer_w_NAs$Cat_Weight = replace(summer_w_NAs$Cat_Weight, summer_w_NAs$Cat_Weight >= 79, 'heavy')
summer_w_NAs$Cat_Weight = replace(summer_w_NAs$Cat_Weight, summer_w_NAs$Cat_Weight<=60, 'light')
summer_w_NAs$Cat_Weight = replace(summer_w_NAs$Cat_Weight, 
                               summer_w_NAs$Cat_Weight > 60 & summer_w_NAs$Cat_Weight < 79, 'avg')



cont_age_medal_w_na <- table(summer_w_NAs$Cat_Age, summer_w_NAs$Medal)
chitest4 = chisq.test(cont_age_medal_w_na)
chitest4


cont_height_medal_w_na <- table(summer_w_NAs$Cat_Height, summer_w_NAs$Medal)
chitest5 = chisq.test(cont_height_medal_w_na)
chitest5


cont_weight_medal_w_na <- table(summer_w_NAs$Cat_Weight, summer_w_NAs$Medal)
chitest6 = chisq.test(cont_weight_medal_w_na)
chitest6

```


## Chapter 4.2: Does the average age vary between men and women?

Performed a 2-sample t-test to determine whether the mean age of men significantly differs from that of women. Null hypothesis is that the two means do not differ. With p-value < 0.05, we reject the null hypothesis and conclude that the average age of men and women significantly differs in summer Olympic athletes.

```{r ttest, echo=FALSE, include = TRUE}
# 2-Sample T-Test
summer_w_NAs = na.omit(summer_w_NAs, Age)
m=subset(summer_w_NAs, Sex=='M')
f=subset(summer_w_NAs, Sex=='F')
mean(m$Age)
mean(f$Age)

t = t.test(m$Age,f$Age, alternative = "two.sided", var.equal = FALSE)
t
```

## Chapter 4.3: Does age, height, and weight significantly change over the years?

Performed ANOVA test to determine whether the mean age, height, and weight of both male and female athletes separately changes over the 120 years of data (30 Olympic Games). Null hypotheses state that the average age, height, and weight of both males and females does not change over 30 Olympic Games. All of our p-values are < 0.05 meaning we reject the null hypotheses and conclude that the average age, height, and weight of both males and females significantly differs for at least one of the summer Olympic games. 

### Chapter 4.3.1: Height 

```{r ANOVA Height, echo=FALSE, include = TRUE}
# ANOVA - does mean height statistically change over the years
summer_w_NAs = na.omit(summer_w_NAs, Age)
summer_w_NAs$Year = as.factor(summer_w_NAs$Year)
male_aov = summer_w_NAs[summer_w_NAs$Sex == 'M',]
female_aov = summer_w_NAs[summer_w_NAs$Sex == 'F',]
library(RColorBrewer)
# ANOVA for Height
# Male
aov_m_height = aov(Height~Year, data=male_aov)
summary(aov_m_height)
# Female
aov_f_height = aov(Height~Year, data=female_aov)
summary(aov_f_height)

plot(Height~Year, data=male_aov, col = terrain.colors(29), main = 'Mean Male Height per Year', ylim=c(130,220))
plot(Height~Year, data=female_aov, col = terrain.colors(29), main = 'Mean Female Height per Year', ylim=c(130,220))
```

The graphs bellow show a slight increase in height over time for male and female. The graphs also show a larger distribution of heights as time moves on. This is likely to be due to an increase in participation as well as an increase in countries and diversity. 

### Chapter 4.3.2: Age

```{r ANOVA Age, echo=FALSE, include = TRUE}
# ANOVA for Age
aov_m_age = aov(Age~Year, data=male_aov)
summary(aov_m_age)
# Female
aov_f_age = aov(Age~Year, data=female_aov)
summary(aov_f_age)

plot(Age~Year, data=male_aov, col = terrain.colors(29), main = 'Mean Male Age per Year')
plot(Age~Year, data=female_aov, col = terrain.colors(29), main = 'Mean Female Age per Year')

```

The graphs above show a relatively constant age for men among every Olympic game, however, for females we can see a significant increase in age since their initial participation in 1900.Again we can see an increase in distribution of ages just like our height statistic. Although the reasoning for this increase in distribution may be due to an increase in life expectancy and an increase in overall health of a population.

### Chapter 4.3.3: Weight

```{r ANOVA Weight, echo=FALSE, include = TRUE}

# ANOVA for Weight
aov_m_weight = aov(Weight~Year, data=male_aov)
summary(aov_m_weight)
# Female
aov_f_weight = aov(Weight~Year, data=female_aov)
summary(aov_f_weight)

plot(Weight~Year, data=male_aov, col = terrain.colors(29), main = 'Mean Male Weight per Year', ylim = c(20,150))
plot(Weight~Year, data=female_aov, col = terrain.colors(29), main = 'Mean Female Weight per Year', ylim = c(20,150))
```

In the graphs above we can see that weight has stayed relatively constant for both male and females for all Olympic games, although there may be a slight increase over time. Like every other statistics the distribution gets larger as time move forward.

```{r tukeyHSD, echo= FALSE, include=FALSE}
tukey_f_height <- TukeyHSD(aov_f_height)
tukey_f_height
tukey_m_height <- TukeyHSD(aov_m_height)
tukey_m_height
tukey_f_age <- TukeyHSD(aov_f_age)
tukey_f_age
tukey_m_age <- TukeyHSD(aov_m_age)
tukey_m_age
tukey_f_weight <- TukeyHSD(aov_f_weight)
tukey_f_weight
tukey_m_weight <- TukeyHSD(aov_m_height)
tukey_m_weight
```


# Chapter 5: Conclusion

## Chapter 5.1: Additional Information 

The dataset also contains the name of the cities that hosted the events; however, it does not have the name of the countries that hosted them. In our opinion, having the names of the hosting countries would have been beneficial because we could have seen what are the countries that hosted the largest number of events over time, and we could have also tested whether the athletes competing at the events hosted by their own country have an advantage over the other athletes. We envision that further analysis would include additional datasets that link city to country. 

## Chapter 5.2: Question Changes?

Most of our SMART questions did not change after the exploratory data analysis as we had all the data we needed to find an answer them. However, our EDA did bring about new SMART questions which would be worth exploring more in depth. For example, by conducting Chi Square tests we found out that age, weight and eight do have an effect on whether an athlete earns a medal, but we did not analyze in what way those variables have an effect on the outcome and if there are some specific sports where these variables matters the most. Therefore, from the initial EDA we got new questions that could and should be answered through another EDA of the dataset. 

## Chapter 5.3: Summary

In conclusion our basic EDA has revealed several interesting statistics in regard to the Olympic games and the athletes participating in them. First and foremost, that the number of athletes have been steadily increasing up until 2008 in which we have begun to see a slight decline in participation. We have also seen female participation increasing in recent years while male participation has been declining in the same time period. When it comes to medals won we can see that the United States is in a comfortable lead for every medal relatively to its competitors. Russia and the USSR follow closely behind while Germany and Australia take 3rd and 4th place respectively in medal count. We have also revealed in our analysis that age, height and weight do influence whether an athlete wins a medal or not. Additional discoveries that were made during our analysis include the differences in age among males and females, as well as, the average age, weight and height are not constant from one Olympic game to the next.

## Chapter 5.4: Future Direction

Moving forward we believe a more in-depth EDA and more correlative SMART questions are possible. As mentioned previously we would like to explore how age, height and weight affect the medal awarded. That is how do each of these statistics correlate to the type of medal won. We would also break this analysis down by competition since weight and height play different rolls in their respective sports. In addition to this medal distribution as affected by athlete statistics we could also investigate how the hosting country affects the medal distribution by country. Does the host win more medals than they normally do when in a non-hosting year? Also how does hosting the Olympics affect the host countries performance in subsequent Olympic games.

# Bibliography

Griffin, R. (2019). 120 years of Olympic history: athletes and results. [online] Kaggle.com. Available at: https://www.kaggle.com/heesoo37/120-years-of-olympic-history-athletes-and-results [Accessed May 2018].
